---
title: "ProblemSet2"
author: "Yiming Chen"
format: pdf
editor: visual
---

## Problem 1 - Modified Random walk

### a

```{r}
random_walk1 <- function(n, seed = NULL) {
  set.seed(seed)
  directions <- sample(c(-1, 1), n, replace = TRUE)
  probs <- runif(n)

  steps <- numeric(n)
  for (i in 1:n) {
    if (directions[i] == 1) {
      steps[i] <- ifelse(probs[i] < 0.05, 10, 1)
    } else {
      steps[i] <- ifelse(probs[i] < 0.20, -3, -1)
    }
  }
  sum(steps)
}
```

```{r}
random_walk2 <- function(n, seed = NULL) {
  set.seed(seed)
  directions <- sample(c(-1, 1), n, replace = TRUE)
  probs <- runif(n)

  steps <- ifelse(
    directions == 1,
    ifelse(probs < 0.05, 10, 1),
    ifelse(probs < 0.20, -3, -1)
  )
  sum(steps)
}
```

```{r}
random_walk3 <- function(n, seed = NULL) {
  set.seed(seed)
  directions <- sample(c(-1, 1), n, replace = TRUE)
  probs <- runif(n)
  
  steps <- sapply(1:n, function(i) {
    if (directions[i] == 1) {
      ifelse(probs[i] < 0.05, 10, 1)
    } else {
      ifelse(probs[i] < 0.20, -3, -1)
    }
  })
  sum(steps)
}
```

```{r}
random_walk1(10)
random_walk2(10)
random_walk3(10)
random_walk1(1000)
random_walk2(1000)
random_walk3(1000)
```

### b

```{r}
random_walk1(10, seed = 42)
random_walk2(10, seed = 42)
random_walk3(10, seed = 42)
random_walk1(1000, seed = 42)
random_walk2(1000, seed = 42)
random_walk3(1000, seed = 42)
```

### c

```{r}
library(microbenchmark)
n <- 1000
microbenchmark(
  loop = random_walk1(n, seed = 42),
  vectorized = random_walk2(n, seed = 42),
  applyfun = random_walk3(n, seed = 42),
  times = 50
)
n <- 100000
microbenchmark(
  loop = random_walk1(n, seed = 42),
  vectorized = random_walk2(n, seed = 42),
  applyfun = random_walk3(n, seed = 42),
  times = 20
)
```

The vectorized implementation (`random_walk2`) consistently outperforms both alternatives, demonstrating the efficiency of R’s vectorized operations. In contrast, the loop-based approach (`random_walk1`) scales poorly but still surpasses the apply-based version. The apply implementation (`random_walk3`) proves to be the slowest for both small and large inputs.

### d

```{r}
estimate_prob1 <- function(n, trials = 100000) {
  results <- replicate(trials, random_walk1(n))
  mean(results == 0)
}
estimate_prob2 <- function(n, trials = 100000) {
  results <- replicate(trials, random_walk2(n))
  mean(results == 0)
}
estimate_prob3 <- function(n, trials = 100000) {
  results <- replicate(trials, random_walk3(n))
  mean(results == 0)
}
set.seed(42)

prob2_10   <- estimate_prob2(10, trials = 10000)
prob2_100  <- estimate_prob2(100, trials = 10000)
prob2_1000 <- estimate_prob2(1000, trials = 10000)
cat("The probability that the random walk ends at 0 with 10 steps is:", prob2_10, '\n')
cat("The probability that the random walk ends at 0 with 100 steps is:", prob2_100, '\n')
cat("The probability that the random walk ends at 0 with 1000 steps is:", prob2_1000)
```

## Problem 2 - Mean of Mixture of Distributions

```{r}
estimate_daily_mean <- function(trials = 100000, seed = 42) {
  set.seed(seed)
  counts <- rpois(trials, 8) + 2 * rnorm(trials, mean = 60, sd = sqrt(12)) + rpois(trials, 64) + rpois(trials, 72)
  return(counts)
}


n <- estimate_daily_mean(trials = 200000, seed = 42)
cat('The estimation of the average number of cars \n that pass an intersection per day under assumptions is:', round(mean(n)))
```

## Problem 3 - Linear Regression

### a

```{r}
youtube <- read.csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-03-02/youtube.csv')
colnames(youtube)
```

```{r}
youtube_deid <- youtube[, c("year","funny","show_product_quickly","patriotic",
                            "celebrity","danger","animals","use_sex",
                            "view_count","like_count","dislike_count",
                            "favorite_count","comment_count","category_id")]
dim(youtube_deid)

```

The dimensions of the data after removing these columns are 247 \* 14 (247 rows \* 14 columns).

### b

```{r}
hist(youtube_deid$view_count, main="view_count", xlab="")
hist(youtube_deid$like_count, main="like_count", xlab="")
hist(youtube_deid$dislike_count, main="dislike_count", xlab="")
hist(youtube_deid$favorite_count, main="favorite_count", xlab="")
hist(youtube_deid$comment_count, main="comment_count", xlab="")
```

For these five variables, favorite_count only has 0 and null values, so it would not be appropriate to use as the outcome in a linear regression model. All the rest variables are right-skewed,so we can carry out the log(1+x) transformation to make it prior to being used as the outcome in a linear regression model. The histograms after transformation are shown below.

```{r}
hist(log1p(youtube_deid$view_count), main="view_count", xlab="")
hist(log1p(youtube_deid$like_count), main="like_count", xlab="")
hist(log1p(youtube_deid$dislike_count), main="dislike_count", xlab="")
hist(log1p(youtube_deid$comment_count), main="comment_count", xlab="")
```

### c

```{r}
Sys.setenv(LANG = "en")
Sys.setlocale("LC_ALL", "C")
youtube_deid$log_views <- log1p(youtube$view_count)
youtube_deid$log_likes <- log1p(youtube$like_count)
youtube_deid$log_dislikes <- log1p(youtube$dislike_count)
youtube_deid$log_comments <- log1p(youtube$comment_count)
model_views <- lm(log_views ~ funny + show_product_quickly + patriotic +
                    celebrity + danger + animals + use_sex + year,
                  data = youtube_deid)

model_likes <- lm(log_likes ~ funny + show_product_quickly + patriotic +
                    celebrity + danger + animals + use_sex + year,
                  data = youtube_deid)

model_dislikes <- lm(log_dislikes ~ funny + show_product_quickly + patriotic +
                       celebrity + danger + animals + use_sex + year,
                     data = youtube_deid)

model_comments <- lm(log_comments ~ funny + show_product_quickly + patriotic +
                       celebrity + danger + animals + use_sex + year,
                     data = youtube_deid)
```

```{r}
summary(model_views)
```

For the model of view_count, none of the attributes shows statistically significant with the view counts.

```{r}
summary(model_likes)
```

For the model of like_count, only the attribute---'year' shows statistically significant association which is also positive. And the 'danger' attribute tends to have a positive and statistically significant association.

```{r}
summary(model_dislikes)
```

For the model of dislike_count, same as like_count, only 'year' shows a positive statistically significant association and the patriotic attribute tends to have a positive statistically significant association with dislike counts.

```{r}
summary(model_comments)
```

For comment_count, only 'year' and 'patriotic' show a tendency of positive significant association and other attributes show no statistically significant association.

### d

```{r}
dat <- na.omit(youtube_deid[, c("log_views","funny","show_product_quickly",
                                "patriotic","celebrity","danger","animals","use_sex","year")])

y <- dat$log_views
X <- model.matrix(~ funny + show_product_quickly + patriotic +
                    celebrity + danger + animals + use_sex + year,
                  data = dat)

beta_hat <- solve(t(X) %*% X) %*% (t(X) %*% y)

coef(model_views)

beta_hat
```

The two results are the same.
